Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-02-23 doc/manuscript/ESSDmanuscript.Rmd
worked on abstract & intro
@mawilliams99 , FYI:

- first pass at abstract
- added a bit of content to intro

---
title: Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'essd')`"
author:
  - given_name: Madison 
    surname: Williams
    affiliation: "1"
  - given_name: Valentine
    surname: Herrmann
    affiliation: "1"
  - given_name: Rebecca
    surname: Banbury Morgan
    affiliation: "1,2" 
  - given_name: Ben 
    surname: Bond-Lamberty
    affiliation: "3" 
  - given_name: Susan 
    surname: Cook-Patton
    affiliation: "4" 
  - given_name: Helene
    surname: Muller-Landau
    affiliation: "5" 
  - given_name: Camille
    surname: Piponiot
    affiliation: "6" 
  - given_name: Teagan
    surname: Rogers
    affiliation: "1"
  - given_name: Kristina J.
    surname: Anderson-Teixeira
    affiliation: "1,5 *"
    email: teixeirak@si.edu
    corresponding: true
affiliation:
  - code: 1
    address: Center for Conservatiton Ecology, Smithsonian Conservation Biology Institute, Front Royal, VA, United States
  - code: 2
    address: School of Geography, University of Leeds, Leeds, UK
  - code: 3
    address: Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD, United States
  - code: 4
    address: The Nature Conservancy; Arlington VA 22203, USA
  - code: 5
    address: Forest Global Earth Observatory, Smithsonian Tropical Research Institute, Panama, Republic of Panama
  - code: 6
    address: CIRAD, Montpellier, France
abstract: |
  The abstract goes here.
  It can also be on _multiple lines_.
bibliography: references.bib
running:
  title: IPCC forest C accounting with ForC
  author: Williams et al.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:
algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company. 
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #  use this to add a statement when having only software code available
  #  use this to add a statement when having only data sets available
  #code: |
  #data: |
  codedata: |
    use this to add a statement when having data sets and software code available
authorcontribution: |
  (fill this in)
disclaimer: |
acknowledgements: |
  Thank you to all researchers who collected and published the data contained in ForC, and to all research assistants and collaborators who have helped to build the database.
  Funding for this study was provided by Bezos Earth Fund to The Nature Conservancy, the Institute for Global Environmental Strategies, WLS(?)
appendix: |
output:
  rticles::copernicus_article: default
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
---

**Important**: Always double-check with the official manuscript preparation guidelines at [https://publications.copernicus.org/for_authors/manuscript_preparation.html](https://publications.copernicus.org/for_authors/manuscript_preparation.html), especially the sections "Technical instructions for LaTeX" and "Manuscript composition".
Please contact Daniel Nüst, `daniel.nuest@uni-muenster.de`, with any problems.

@@

Forests are critical to our management of greenhouse gases. The influence the forests have on asmosphereic carbon has made forest data increasingly valuable.

@@

>>>>

---
title: Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'essd')`"
author:
  - given_name: Madison 
    surname: Williams
    affiliation: "1"
  - given_name: Valentine
    surname: Herrmann
    affiliation: "1"
  - given_name: Rebecca
    surname: Banbury Morgan
    affiliation: "1,2" 
  - given_name: Ben 
    surname: Bond-Lamberty
    affiliation: "3" 
  - given_name: Susan 
    surname: Cook-Patton
    affiliation: "4" 
  - given_name: Helene
    surname: Muller-Landau
    affiliation: "5" 
  - given_name: Camille
    surname: Piponiot
    affiliation: "6" 
  - given_name: Teagan
    surname: Rogers
    affiliation: "1"
  - given_name: Kristina J.
    surname: Anderson-Teixeira
    affiliation: "1,5 *"
    email: teixeirak@si.edu
    corresponding: true
affiliation:
  - code: 1
    address: Center for Conservatiton Ecology, Smithsonian Conservation Biology Institute, Front Royal, VA, United States
  - code: 2
    address: School of Geography, University of Leeds, Leeds, UK
  - code: 3
    address: Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD, United States
  - code: 4
    address: The Nature Conservancy; Arlington VA 22203, USA
  - code: 5
    address: Forest Global Earth Observatory, Smithsonian Tropical Research Institute, Panama, Republic of Panama
  - code: 6
    address: CIRAD, Montpellier, France
abstract: |
  Forests are critical for climate change mitigation and consitute a substantial portion of planned emissions reductions under the 2015 Paris Agreement.
  Yet, the efficacy of greenhouse gas (GHG) mitigation planning and reporting is dependent upon the quality of available emission factors data, including forest carbon (C) stocks and changes therein.
  Tens of thousands of relevant forest C estimates have been published, yet are not readily accesible to the practitioners compiling national greenhouse gas inventories. 
  Many of these data have, however, been compiled in the Global Forest C database (ForC; https://forc-db.github.io/) and stand to be of value to greenhouse gas accounting if made available through the Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC).
  Here, we develop and document a process for semi-automated transfer of data from ForC into the EFDB, assess the data available and transferred to date, and provide recommendations for improving forest data collection, analysis, and reporting to improve accounting of forest-sector GHG emissions and removals.
  We begin by reconciling terminology and mapping ForC fields into EFDB.
  This process required some updates to the ForC database structure, leading to the release of a new version of ForC (v4.0; described here).
  At the time of writing, ForC contained ## values that would qualify for inclusion in the EFDB, ## of which have been transferred to date.
  (Some analysis of representation/ gaps)
  <!-- -->
  In the future, forest C estimates in EFDB can be improved through targetted research to fill critical gaps, reporting of information required by IPCC, and continued submission of data from scientific publications to the EFDB.

bibliography: references.bib
running:
  title: IPCC forest C accounting with ForC
  author: Williams et al.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:
algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company. 
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #  use this to add a statement when having only software code available
  #  use this to add a statement when having only data sets available
  #code: |
  #data: |
  codedata: |
    use this to add a statement when having data sets and software code available
authorcontribution: |
  (fill this in)
disclaimer: |
acknowledgements: |
  Thank you to all researchers who collected and published the data contained in ForC, and to all research assistants and collaborators who have helped to build the database.
  Funding for this study was provided by Bezos Earth Fund to The Nature Conservancy, the Institute for Global Environmental Strategies, WLS(?)
appendix: |
output:
  rticles::copernicus_article: default
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
---

<!--**Important**: Always double-check with the official manuscript preparation guidelines at [https://publications.copernicus.org/for_authors/manuscript_preparation.html](https://publications.copernicus.org/for_authors/manuscript_preparation.html), especially the sections "Technical instructions for LaTeX" and "Manuscript composition".
Please contact Daniel Nüst, `daniel.nuest@uni-muenster.de`, with any problems.-->

@@

Forests are critical to our management of greenhouse gases. 
The influence the forests have on atmospheric carbon has made forest data increasingly valuable.

**Forests play a central role in international plans for climate change mitigation [@unfccc_adoption_2015].**
As of 2016, forest-based mitigation accounted for 26% of total planned greenhouse gas mitigation within Nationally Determined Contributions under the Paris Agreement [@grassi_key_2017].
Effective forest-based climate change mitigation requires solid scientific data [@anderson-teixeira_effective_2022].

@@

**To ensure that planned emissions reductions are realistic, high-quality estimates of forest C stocks and fluxes must be publicly accessible.**
The EFDB is intended as a recognized library, and can be used both for efforts to tally a nation's intended or accomplished GHG reductions, or as a basis of comparison for external parties to evaluate these inventories. 
<!-- check Washington Post article on Malaysia NDC, @deng_comparing_2021 -->

### 2022-02-23 doc/manuscript/ESSDmanuscript.tex
worked on abstract & intro
@mawilliams99 , FYI:

- first pass at abstract
- added a bit of content to intro

%% Journal abbreviations (please use the same for discussion papers and final revised papers)

@@

\begin{abstract}
The abstract goes here. It can also be on \emph{multiple lines}.
\end{abstract}

@@

\textbf{Important}: Always double-check with the official manuscript
preparation guidelines at
\url{https://publications.copernicus.org/for_authors/manuscript_preparation.html},
especially the sections ``Technical instructions for LaTeX'' and
``Manuscript composition''. Please contact Daniel Nüst,
\texttt{daniel.nuest@uni-muenster.de}, with any problems.

@@

\textbf{As we seek to mitigate climate change
\citep{unfccc_adoption_2015}, forests are critical.} \emph{(See intros
to other papers for relevant refs)} Under Paris Agreement,
\textbf{\#\#}\% of Nationally Determined Contributions relate to forests
\citep{grassi_key_2017}.

\textbf{The International Panel on Climate Change (IPCC) provides
guidance for national greenhouse gas (GHG) inventories for reporting to
the United Nations Framework Convention on Climate Change \citep[UNFCCC,
(REFS for older guidelines),][]{ipcc_2019_2019}.} Under this guidance,
GHG inventories include all managed land, including most of the world's
forest land \citep{ogle_delineating_2018}. The IPCC inventory guidelines
include specific instructions for accounting for forest land \ldots{}
\citep{ipcc_agriculture_2006, ipcc_2019_2019a}. This guidance has
improved as more of the relevant underlying data has become available
\citep{requenasuarez_estimating_2019}, but there remains room for
continuous improvement as the science advances. For example,
\citet{cook-patton_mapping_2020} \emph{(summarize IPCC comparison
results)}. \emph{Moreover, it is useful for those compiling national
greenhouse gas inventories to have access to locally-specific
information, when available.} To improve the data accessible for C
accounting, the IPCC created the Emission Factor Database (EFDB;
https://www.ipcc-nggip.iges.or.jp/EFDB/main.php), which is intended as a
recognized library of emission factors and other parameters that can be
used for estimating greenhouse gas emissions and removals.

>>>>

%% Journal abbreviations (please use the same for preprints and final revised papers)

@@

\begin{abstract}
Forests are critical for climate change mitigation and consitute a
substantial portion of planned emissions reductions under the 2015 Paris
Agreement. Yet, the efficacy of greenhouse gas (GHG) mitigation planning
and reporting is dependent upon the quality of available emission
factors data, including forest carbon (C) stocks and changes therein.
Tens of thousands of relevant forest C estimates have been published,
yet are not readily accesible to the practitioners compiling national
greenhouse gas inventories. Many of these data have, however, been
compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector GHG emissions and
removals. We begin by reconciling terminology and mapping ForC fields
into EFDB. This process required some updates to the ForC database
structure, leading to the release of a new version of ForC (v4.0;
described here). At the time of writing, ForC contained \#\# values that
would qualify for inclusion in the EFDB, \#\# of which have been
transferred to date. (Some analysis of representation/ gaps) In the
future, forest C estimates in EFDB can be improved through targetted
research to fill critical gaps, reporting of information required by
IPCC, and continued submission of data from scientific publications to
the EFDB.
\end{abstract}

@@

@@

Forests are critical to our management of greenhouse gases. The
influence the forests have on atmospheric carbon has made forest data
increasingly valuable.

\textbf{Forests play a central role in international plans for climate
change mitigation \citep{unfccc_adoption_2015}.} As of 2016,
forest-based mitigation accounted for 26\% of total planned greenhouse
gas mitigation within Nationally Determined Contributions under the
Paris Agreement \citep{grassi_key_2017}. Effective forest-based climate
change mitigation requires solid scientific data
\citep{anderson-teixeira_effective_2022}.

\textbf{The International Panel on Climate Change (IPCC) provides
guidance for national greenhouse gas (GHG) inventories for reporting to
the United Nations Framework Convention on Climate Change \citep[UNFCCC,
(REFS for older guidelines),][]{ipcc_2019_2019}.} Under this guidance,
GHG inventories include all managed land, including most of the world's
forest land \citep{ogle_delineating_2018}. The IPCC inventory guidelines
include specific instructions for accounting for forest land \ldots{}
\citep{ipcc_agriculture_2006, ipcc_2019_2019a}. This guidance has
improved as more of the relevant underlying data has become available
\citep{requenasuarez_estimating_2019}, but there remains room for
continuous improvement as the science advances. For example,
\citet{cook-patton_mapping_2020} \emph{(summarize IPCC comparison
results)}. Another example is \citet{cuni-sanchez_high_2021}.
\emph{Moreover, it is useful for those compiling national greenhouse gas
inventories to have access to locally-specific information, when
available.} To improve the data accessible for C accounting, the IPCC
created the Emission Factor Database (EFDB;
https://www.ipcc-nggip.iges.or.jp/EFDB/main.php), which is intended as a
recognized library of emission factors and other parameters that can be
used for estimating greenhouse gas emissions and removals.

\textbf{To ensure that planned emissions reductions are realistic,
high-quality estimates of forest C stocks and fluxes must be publicly
accessible.} The EFDB is intended as a recognized library, and can be
used both for efforts to tally a nation's intended or accomplished GHG
reductions, or as a basis of comparison for external parties to evaluate
these inventories.
